---
title: "Combined_Stats2_Project1_AC_AP_GL"
author: "Adam Canton, Anish Patel & Grace Lang"
date: "6/7/2020"
output: word_document
---

Kaggle link for data reference: https://www.kaggle.com/kumarajarshi/life-expectancy-who

```{r setup, include=FALSE}

library(class)
library(caret)
library(e1071)
library(magrittr)
library(dplyr)
library(tidyr)
library(naniar)
library(ggplot2)
library(plotly)
library(forcats)
library(ggExtra)
library(glmnet)
library(GGally)
library(ggcorrplot)
library(car)
library(olsrr)
library(rgl)
library(tree)
library(ISLR)
library(leaps)
library(matrixStats)

#Adam's Connections
#life <- read.csv(file = "F:/R For Real/Stats2 Project 1/Life Expectancy Data.csv")

#Anish's Connections
#life <- read.csv("Life Expectancy Data.csv")

#Grace's Connections
life <-read.csv("C:/Users/david/OneDrive/Desktop/AppliedStats/6372_AppliedStats_GraceLang/Project1/LifeExpectancyData.csv")
#life <-read.csv("C:/Users/Dave/Desktop/AppliedStats/Stats2_Project1/LifeExpectancyData.csv")


#Just pull out 2014 into the data
Life14 <- subset(life, Year==2014)
summary(Life14)
```
#Column definitions:
Adult.mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)
Infant.death: Number of Infant Deaths per 1000 population
Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)
percentage expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita(%)
Hepatitis B: Hepatitis B (HepB) immunization coverage among 1-year-olds (%)
Measles - number of reported cases per 1000 population
BMI: Average Body Mass Index of entire population
Under-five-deaths: Number of under-five deaths per 1000 population
Polio: Polio (Pol3) immunization coverage among 1-year-olds (%)
Total expenditure: General government expenditure on health as a percentage of total government expenditure (%)
Diphtheria: Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds (%)
HIV/AIDS: Deaths per 1 000 live births HIV/AIDS (0-4 years)
GDP: Gross Domestic Product per capita (in USD)
Population: Population of the country
thinness: Prevalence of thinness among children and adolescents for Age 10 to 19 (% )
Thinness 5-9: Prevalence of thinness among children for Age 5 to 9(%)
income composition: Human Development Index in terms of income composition of resources (index ranging from 0 to 1)
Schooling: Number of years of Schooling(years)


##Things to note from summary stats:
 * 83% of data comes from developing countries, life exp may be lower.
 * Mean life expectancy is 71.54 yr old
 * The max for Measles seems outrageously high 79k - which country? outlier?
 * Looks like there's a couple of positive correlations with life expectancy: income.compostion & schooling
 * Adult mortality seems to have a gender split, but cannot determine for certain
 * No outward evidence of any of the parameters having a non-linear relationship (adding a quadratic)
 
 # Fixing NA's in GDP
```{r}
Life14$GDP[159] = 2520.2027
Life14$GDP[11] = 29436.13
Life14$GDP[20] = 3081.23
Life14$GDP[28] = 1559.38
Life14$GDP[39] = 6208.58
Life14$GDP[44] = 19771.65
Life14$GDP[45] = 1568.63
Life14$GDP[46] = 352.82
Life14$GDP[51] = 3378.68
Life14$GDP[54] = 1209.68
Life14$GDP[61] = 607.21
Life14$GDP[77] = 5608.62
Life14$GDP[89] = 1279.64
Life14$GDP[90] = 1998.49
Life14$GDP[95] = 6466.52
Life14$GDP[107] = 2971.73
Life14$GDP[131] = 27802.96
Life14$GDP[132] = 2674.35
Life14$GDP[136] = 8743.89
Life14$GDP[137] = 6684.67
Life14$GDP[146] = 18671.59
Life14$GDP[162] = 5487.92
Life14$GDP[173] = 47614.61
Life14$GDP[174] = 1000.00
Life14$GDP[175] = 55025.13
Life14$GDP[179] = 16053.24
Life14$GDP[180] = 2030.31
Life14$GDP[181] = 1674.28
```
# Fixing Population NA's
```{r}
Life14$Population[159] = 18715672
Life14$Population[11] = 370633
Life14$Population[20] = 10710000
Life14$Population[28] = 22650000
Life14$Population[39] = 5244359
Life14$Population[44] = 10510000
Life14$Population[45] = 25500000
Life14$Population[46] = 10780263
Life14$Population[51] = 90420000
Life14$Population[54] = 5580000
Life14$Population[61] = 2024000
Life14$Population[77] = 77470000
Life14$Population[89] = 5836000
Life14$Population[90] = 6640000
Life14$Population[95] = 6362000
Life14$Population[107] = 107446
Life14$Population[131] = 50750000
Life14$Population[132] = 3556000
Life14$Population[136] = 178296
Life14$Population[137] = 108861
Life14$Population[146] = 5420000
Life14$Population[162] = 2070000
Life14$Population[173] = 64350000
Life14$Population[174] = 49960000
Life14$Population[175] = 318400000
Life14$Population[179] = 30050000
Life14$Population[180] = 91710000
Life14$Population[181] = 25820000
Life14$Population[130] = 2459000
Life14$Population[145] = 5470000
Life14$Population[172] = 9214000
Life14$Population[88] = 3691000
Life14$Population[12] = 1336000
Life14$Population[140] = 30920000
Life14$Population[14] = 284825
Life14$Population[5] = 92562
Life14$Population[66] = 108902
Life14$Population[116] = 4510000
Life14$Population[24] = 409769
Life14$Population[121] = 4027000
Life14$Population[149] = 13420000
Life14$Population[42] = 11310000
```
# Fixing 27 - 0 Percent Expenditures - Likely going to turn this variable into a categorical with 3-5 levels depending on spread
```{r}
Life14$percentage.expenditure[11] = 748.36
Life14$percentage.expenditure[20] = 115.35
Life14$percentage.expenditure[28] = 16.85
Life14$percentage.expenditure[39] = 34.78
Life14$percentage.expenditure[44] = 1249.23
Life14$percentage.expenditure[46] = 2.68
Life14$percentage.expenditure[51] = 48.66
Life14$percentage.expenditure[54] = 2.82
Life14$percentage.expenditure[61] = 4.93
Life14$percentage.expenditure[77] = 192.78
Life14$percentage.expenditure[89] = 35.5
Life14$percentage.expenditure[90] = 13.79
Life14$percentage.expenditure[95] = 197.84
Life14$percentage.expenditure[107] = 102.43
Life14$percentage.expenditure[131] = 1093.43
Life14$percentage.expenditure[132] = 113.87
Life14$percentage.expenditure[136] = 189.84
Life14$percentage.expenditure[137] = 178.51
Life14$percentage.expenditure[146] = 1024.94
Life14$percentage.expenditure[159] = 31.65
Life14$percentage.expenditure[162] = 219.74
Life14$percentage.expenditure[173] = 3682.48
Life14$percentage.expenditure[174] = 11.47
Life14$percentage.expenditure[175] = 4541.8
Life14$percentage.expenditure[179] = 235.35
Life14$percentage.expenditure[180] = 48.82
Life14$percentage.expenditure[181] = 12.9
```
# Fixing the 10 missing school variables - well except somalia (setting to 0) - might considering leaving somlia out it has almost no data
```{r}
Life14$Schooling[28] = 5.2
Life14$Schooling[44] = 12.7
Life14$Schooling[45] = 10.8
Life14$Schooling[46] = 6.8
Life14$Schooling[131] = 12.2
Life14$Schooling[132] = 11.6
Life14$Schooling[149] = 0
Life14$Schooling[173] = 17.4
Life14$Schooling[174] = 8
Life14$Schooling[175] = 16.3
```
#What variables are still missing
```{r replacing NAs with Mean}
gg_miss_var(Life14)

#Can I add the avgs into the NAs without a lot of shift in data
#Life14 %>% ggplot(aes(x = Income.composition.of.resources)) + geom_histogram() #normally distributed, low effect 2% change in mean
#Life14 %>% ggplot(aes(x = Hepatitis.B)) + geom_histogram() #left-skewed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = Total.expenditure)) + geom_histogram() #normally distributed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = thinness.5.9.years)) + geom_histogram() #right-skewerd, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = thinness..1.19.years)) + geom_histogram() #right-skewerd, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = BMI)) + geom_histogram() #normally distributed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = Alcohol)) + geom_histogram() #right-skewerd, low effect 0% change in mean

#Replacing NAs with the mean of the column
Life14$Income.composition.of.resources[is.na(Life14$Income.composition.of.resources)] <- round(mean(Life14$Income.composition.of.resources, na.rm = TRUE))
Life14$Hepatitis.B[is.na(Life14$Hepatitis.B)] <- round(mean(Life14$Hepatitis.B, na.rm = TRUE))
Life14$Total.expenditure[is.na(Life14$Total.expenditure)] <- round(mean(Life14$Total.expenditure, na.rm = TRUE))
Life14$thinness..1.19.years[is.na(Life14$thinness..1.19.years)] <- round(mean(Life14$thinness..1.19.years, na.rm = TRUE))
Life14$thinness.5.9.years[is.na(Life14$thinness.5.9.years)] <- round(mean(Life14$thinness.5.9.years, na.rm = TRUE))
Life14$BMI[is.na(Life14$BMI)] <- round(mean(Life14$BMI, na.rm = TRUE))
Life14$Alcohol[is.na(Life14$Alcohol)] <- round(mean(Life14$Alcohol, na.rm = TRUE))


```
# Adding in some feature engineering
```{r}
#Factorizing things
IncomeCompFactor = cut(Life14$Income.composition.of.resources, breaks = c(0,0.4,0.7,1), labels = c("Low", "Moderate", "High"))

HepBfactor = cut(Life14$Hepatitis.B, breaks = c(0,25,50,75,100), labels = c("Deficient", "Low", "Moderate", "High"))

PercentExpendFactor = cut(Life14$percentage.expenditure, breaks = c(0,5000,15000,20000), labels = c("Low", "Moderate", "High"))

PopulationFactor = cut(Life14$Population, breaks = c(0, 3.612e+05,2.196e+07,1.294e+09), labels = c("Low", "Moderate", "High"))

GdpFactor = cut(Life14$GDP, breaks = c(0,40000,80000,120000), labels = c("Low","Moderate","High"))

SchoolFactor = cut(Life14$Schooling, breaks = c(0,12,16,21), labels = c("Low","Moderate", "High"))

# Binding them back to Life14
Life14 = cbind(Life14, HepBfactor, PercentExpendFactor, PopulationFactor, GdpFactor, SchoolFactor)

#Fixing some errors where for some reason it wasn't picking up these 3
Life14$PercentExpendFactor[45] = as.factor("Low")
Life14$PercentExpendFactor[149] = as.factor("Low")
Life14$SchoolFactor[149] = as.factor("Low")

#adding log variables
Life14 = Life14 %>% mutate(LHIV.AIDS = log(HIV.AIDS))
Life14 = Life14 %>% mutate(Linfant.deaths = log(infant.deaths + 1))
Life14 = Life14 %>% mutate(LUnder5Deaths = log(under.five.deaths + 1))
Life14 = Life14 %>% mutate(LThin19 = log(thinness..1.19.years + 1))
Life14 = Life14 %>% mutate(LGDP = log(GDP))

# Creating scaled variables for knn testing
Life14Scale = Life14
Life14Scale$Life.expectancy = scale(Life14Scale$Life.expectancy) 
Life14Scale$Adult.Mortality = scale(Life14Scale$Adult.Mortality)
Life14Scale$percentage.expenditure = scale(Life14Scale$percentage.expenditure)
Life14Scale$Diphtheria = scale(Life14Scale$Diphtheria)
Life14Scale$Hepatitis.B = scale(Life14Scale$Hepatitis.B)
Life14Scale$Income.composition.of.resources = scale(Life14Scale$Income.composition.of.resources)
Life14Scale$Schooling = scale(Life14Scale$Schooling)
Life14Scale$LHIV.AIDS = scale(Life14Scale$LHIV.AIDS)
Life14Scale$Linfant.deaths = scale(Life14Scale$Linfant.deaths)
Life14Scale$LUnder5Deaths = scale(Life14Scale$LUnder5Deaths)
Life14Scale$LThin19 = scale(Life14Scale$LThin19)
Life14Scale$LGDP = scale(Life14Scale$LGDP)
Life14Scale$Total.expenditure = scale(Life14Scale$Total.expenditure)

```


#Visualizing some of the data

Boxplot of life expectancy by country status:
```{r}
Life14 %>% ggplot(aes(y = Life.expectancy)) + geom_boxplot() + facet_wrap(~Status)
```

Pairwise plots for selected variables:
```{r message = FALSE, warning = FALSE}
df2 <- select(Life14, c(Life.expectancy, Adult.Mortality, Total.expenditure, HIV.AIDS, Income.composition.of.resources, Schooling))
library(GGally)
ggpairs(df2)
```

Correlation between life expectancy and all the numeric independent variables:
```{r}
cor.xy <- cor(Life14 %>% select(-c(Country, Year, Status)), use = "complete.obs")
LE.cor <- data.frame(var = rownames(cor.xy)[-1], correlation = cor.xy[-1, 1])
LE.cor %>% ggplot(aes(x = var, y = correlation)) + geom_col() + coord_flip()
```

# Checking correlations between variables that are missing data and variables I think may make good substitutes and are not missing data.
It seems that Schooling is a good replacement for income.comp though it seems as though income.comp has a slightly better relationship with life expectancy
I think Im going to go with Schooling as the more complete variable
Diphtheria and HepB also seem like good substitutes and diptheria is the more highly correlated with life expectancy going to throw out income.comp, and HepB, along with year, country, status to check correlations in next chunk.
```{r}
corr <- Life14 %>% 
  select("Income.composition.of.resources", "Hepatitis.B", "Diphtheria", "Schooling", "Measles", "Life.expectancy", "GDP", "Polio") %>%
  drop_na()

corr <- round(cor(corr), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE, lab_size = 3, method = "circle",
           colors = c("tomato2", "white", "springgreen3"),
           title = "Correlations of all relevant variables",
           ggtheme = theme_bw())
```
# Excluded factors here have been selected to due high (>0.9) correlations or are not numeric in nature (Status...)
The results here show that we can likely get rid of: a thinness category, either infant deaths or deaths under 5, and gdp or percent.expenditure
```{r}
# http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Correlogram

exclude_factors <- c("Country", "Status", "Year", "Hepatitis.B", "Income.composition.of.resources")

corr <- Life14 %>% select( -all_of(exclude_factors)) %>% drop_na()

corr <- round(cor(corr), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE, lab_size = 3, method = "circle",
           colors = c("tomato2", "white", "springgreen3"),
           title = "Correlations of all relevant variables",
           ggtheme = theme_bw())
```

```{r}
# The variables I threw out are the ones that were missing data or least correlated with life expectancy - 
# Though admittedly sometimes the difference was small
# Throw out - infant deaths, thinness 1-19, and percent expenditure
# no more significant correlations between explanatory variables - time to look at selection

exclude_factors <- c("Country", "Status", "Year", "Hepatitis.B", 
                     "Income.composition.of.resources", "infant.deaths", "thinness..1.19.years", "percentage.expenditure")

corr <- Life14 %>% select( -all_of(exclude_factors)) %>% drop_na()

corr <- round(cor(corr), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE, lab_size = 3, method = "circle",
           colors = c("tomato2", "white", "springgreen3"),
           title = "Correlations of all relevant variables",
           ggtheme = theme_bw())
```


```{r}
# Drop Country and year from sets - drops 2 variables - finalize the set
exclude_factors <- c("Country", "Year")
Life14 <- Life14 %>% select( -all_of(exclude_factors))

# Set training and Testing Sets
set.seed(1234)
index<-sample(1:dim(Life14)[1],90,replace=F)
train<-Life14[index,]
test<-Life14[-index,]
```

```{r}
# Forward Selection - BIC Suggests 5 predictors - Mallow CP suggests 5 
reg.fwd=regsubsets(Life.expectancy~.,data=train,method="forward",nvmax=20)
bics<-summary(reg.fwd)$bic
plot(1:21,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)
```

```{r}
# Adjr2 suggests up to 9
adjr2<-summary(reg.fwd)$adjr2
plot(1:21,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)
```

```{r}
MallowCP <- summary(reg.fwd)$cp
plot(1:21,MallowCP,type="l",ylab="Mallow's CP",xlab="# of predictors")
index<-which(MallowCP==min(MallowCP))
points(index,MallowCP[index],col="red",pch=10)
```

# Looking at all the parameters significance
```{r}
modelgenerator <- lm(Life.expectancy ~ . , data = train)
ols_step_both_aic(modelgenerator)
#ols_step_both_aic(model3)
```

```{r}
# Looking for a good model using Olsrr going to start with stepwise - using p-values as target values
model1 <- lm(Life.expectancy ~ ., data = Life14)
ols_step_both_p(model1)
vif(model1)
```

```{r}
model1 <- lm(Life.expectancy ~ Schooling + Adult.Mortality + HIV.AIDS + GDP + thinness.5.9.years + Diphtheria + Alcohol, data = Life14)
summary(model1)
vif(model1)
```


```{r}
# Checking fit of model 1 - we actually look really good here. We have 2 points that may be problematic #28 Ivory Coast and #143 Seychelle 
# unsure of why this is nothing really stand out about them.
ols_plot_resid_lev(model1)
# nice looking random cloud
ols_plot_resid_fit(model1)
# Here we see the low end looks a bit un-normal but overall is largely normal
ols_plot_resid_qq(model1)
# as we saw in the QQ plot above there is some left skew here but largely looks normal
ols_plot_resid_hist(model1)
```

#Generating 3 models to cross compare smooth & fit
```{r}
### MODEL 1 ###
modeltest1 <- lm(Life.expectancy ~ Adult.Mortality + SchoolFactor + LHIV.AIDS + percentage.expenditure + Total.expenditure + Linfant.deaths , data = train)
#summary(modeltest1)
#vif(modeltest1)
#ols_plot_resid_fit(modeltest1)
#ols_plot_resid_lev(modeltest1)
#ols_plot_resid_qq(modeltest1)
#ols_plot_resid_hist(modeltest1)
#ols_plot_cooksd_bar(modeltest1)
#ols_step_both_aic(model4)

### MODEL 2 ###
modeltest2 <- lm(Life.expectancy ~ SchoolFactor + LHIV.AIDS + Adult.Mortality + percentage.expenditure + 
                   Linfant.deaths + Status + LGDP + Diphtheria , data = train)
#summary(modeltest2)
#vif(modeltest2)
#ols_plot_resid_fit(modeltest2)
#ols_plot_resid_lev(modeltest2)
#ols_plot_resid_qq(modeltest2)
#ols_plot_resid_hist(modeltest2)
#ols_plot_cooksd_bar(modeltest2)

### MODEL 3  - ALL IN WITH INTERACTIONS ###
modeltest3 <- lm(Life.expectancy ~ 
                    Income.composition.of.resources:Schooling + Population + Measles + Hepatitis.B + Diphtheria + Polio + Total.expenditure + infant.deaths +
                    under.five.deaths + log(HIV.AIDS)	+ BMI	+ thinness.5.9.years + thinness..1.19.years +	Alcohol +	percentage.expenditure +	GDP +	Status +
                    Adult.Mortality	+ Income.composition.of.resources +	HIV.AIDS:Income.composition.of.resources + Schooling +
                    percentage.expenditure:Income.composition.of.resources +	thinness..1.19.years:thinness.5.9.years +	HIV.AIDS:GDP +Hepatitis.B:Diphtheria +
                    BMI:Schooling	+ Total.expenditure:GDP +	Adult.Mortality:Total.expenditure +	HIV.AIDS:Schooling +	under.five.deaths:thinness..1.19.years +
                    Population:thinness..1.19.years +	Adult.Mortality:Polio +	Status:Total.expenditure +	Status:GDP + percentage.expenditure:Population +
                    Alcohol:thinness.5.9.years +	infant.deaths:Population +	Alcohol:HIV.AIDS +	HIV.AIDS:Population +	BMI:under.five.deaths +
                    Total.expenditure:thinness.5.9.years +	Total.expenditure:HIV.AIDS +	percentage.expenditure:HIV.AIDS +
                    Diphtheria:Income.composition.of.resources +	Measles:HIV.AIDS +	Measles:Schooling +	Status:Schooling + Status:Measles +
                    Status:Income.composition.of.resources +	Status:thinness.5.9.years	+ Status:Hepatitis.B +	percentage.expenditure:GDP +
                    percentage.expenditure:Schooling +	GDP:Schooling +	Status:percentage.expenditure +	percentage.expenditure:Polio +	
                    Polio:Total.expenditure	 + Status:Polio +	Measles:thinness..1.19.years	+ Population:thinness.5.9.years +	Polio:Diphtheria,
                    data = train)
#summary(modeltest3)
#vif(modeltest3)
#ols_plot_resid_fit(modeltest3)
#ols_plot_resid_lev(modeltest3)
#ols_plot_resid_qq(modeltest3)
#ols_plot_resid_hist(modeltest3)
```

# Getting results of the 3 Models
```{r}
predictions1 <- modeltest1 %>% predict(test)
d1 = data.frame(R2 = R2(predictions1,test$Life.expectancy),RMSE = RMSE(predictions1,test$Life.expectancy), MAE = MAE(predictions1, test$Life.expectancy))

predictions2 <- modeltest2 %>% predict(test)
d2 = data.frame(R2 = R2(predictions2,test$Life.expectancy),RMSE = RMSE(predictions2,test$Life.expectancy), MAE = MAE(predictions2, test$Life.expectancy))

predictions3 <- modeltest3 %>% predict(test)
d3 = data.frame(R2 = R2(predictions3,test$Life.expectancy),RMSE = RMSE(predictions3,test$Life.expectancy), MAE = MAE(predictions3, test$Life.expectancy))

# Need to remark these values
#data.frame(RMSE = c(d1$RMSE, d2$RMSE, d3$RMSE), Model = c("1", "2", "3"))

Pred1 <- data.frame(Value = predictions1, Model = "1")
Pred2 <- data.frame(Value = predictions2, Model = "2")
Pred3 <- data.frame(Value = predictions3, Model = "3")
PredActual <- data.frame(ActualValue = test$Life.expectancy)
PredAll <- rbind(Pred1, Pred2, Pred3)
PredActual <- rbind(PredActual,PredActual,PredActual)
PredAll <- cbind(PredAll, PredActual)

#Final Plot
PredAll %>% ggplot(aes(x = Value, y = ActualValue, fill = Model)) + geom_point(aes(color = Model)) + geom_smooth(formula = y~x)+
  facet_wrap(facets = PredAll$Model)

#Barchart of results
PredAll  %>% mutate(difference = (Value - ActualValue)) %>% group_by(Model) %>% 
  summarise(RMSE = sqrt(sum(difference^2)/nrow(test))) %>%
  ggplot(aes(x = reorder(Model,RMSE), y = RMSE)) + geom_col(aes(fill = Model)) + 
  geom_text(aes(label = round(RMSE, 4), vjust = -0.5)) + xlab("Model #") + ylab("RMSE on Test Set") + ggtitle("Model RMSE's 70/30 Cross Validation")
```

# Repeating RMSE 
```{r}
# Set number of times you would like to repeat the sampling/testing - currently 500
iterations = 1:500

# the initial values for the columns (might not need these now that ive switched to building columns) 
rmse1 = c()
rmse2 = c()
rmse3 = c()

# Start of Loop
for(i in iterations){
  # Resets sample every iteration 
  index<- sample(1:dim(Life14)[1],128,replace=F)
  train<- Life14[index,]
  test<- Life14[-index,]
  
  # the model runs -  feel free to change them up and see what comes out
  modeltest1 <- lm(Life.expectancy ~ Adult.Mortality + SchoolFactor + LHIV.AIDS + percentage.expenditure + Total.expenditure + Linfant.deaths , data = train)
  
  modeltest2 <- lm(Life.expectancy ~ SchoolFactor + LHIV.AIDS + Adult.Mortality + percentage.expenditure + 
                   Linfant.deaths + Status + LGDP + Diphtheria , data = train)
  
  modeltest3 <- lm(Life.expectancy ~ Adult.Mortality, data = train)
  
  # predictors and column building
  predictions1 <- modeltest1 %>% predict(test)
  
  d1 = data.frame(R2 = R2(predictions1,test$Life.expectancy),
                  RMSE = RMSE(predictions1,test$Life.expectancy), MAE = MAE(predictions1, test$Life.expectancy))
  rmse1 = c(rmse1,d1$RMSE)
  
  predictions2 <- modeltest2 %>% predict(test)

  d2 = data.frame(R2 = R2(predictions2,test$Life.expectancy),
                RMSE = RMSE(predictions2,test$Life.expectancy), MAE = MAE(predictions2, test$Life.expectancy))
  rmse2 = c(rmse2, d2$RMSE)
  
  predictions3 <- modeltest3 %>% predict(test)

  d3 = data.frame(R2 = R2(predictions3,test$Life.expectancy),
                RMSE = RMSE(predictions3,test$Life.expectancy), MAE = MAE(predictions3, test$Life.expectancy))
  rmse3 = c(rmse3, d3$RMSE)
  # End for
}
# putting the dataframe together and outputting relevant statistics
Model.Average.RMSE = cbind(rmse1, rmse2, rmse3)
rmsedf = as.data.frame(Model.Average.RMSE)
Means = colMeans(Model.Average.RMSE)
SDs = round(colSds(Model.Average.RMSE), 3)
range1 = max(rmsedf$rmse1) - min(rmsedf$rmse1)
range2 = max(rmsedf$rmse2) - min(rmsedf$rmse2)
range3 = max(rmsedf$rmse3) - min(rmsedf$rmse3)

# Looking at range of each model to determine how variant it is - can also look at stats above
Means
print("Standard Devs")
SDs
print("ranges - measure of variance between train/test shuffles")
print(paste("Model 1 Range" , round(range1, 3)))
print(paste("Model 2 Range" , round(range2, 3)))
print(paste("Model 3 Range" , round(range3, 3)))
cat("Summary: 
    ")
summary(Model.Average.RMSE)


```

```{r}
# Set knn vectors - these are the column #s that will be analyzed 
knnvector1 = c(3,6,12,20)
knnvector2 = c(3,6,13,26:30)
knnvector3 = c(3,6,13,20)

# Start of Loop Variables
iterations = 1:500
i = 0
knnrmse1 = c()
knnrmse2 = c()
knnrmse3 = c()

# Loop start
for(i in iterations){
  # Create train and test sets
  index<-sample(1:dim(Life14Scale)[1],128,replace=F)
  train<-Life14Scale[index,]
  test<-Life14Scale[-index,]


  #KNN1
  predknn1 <- knn.reg(train[,knnvector1], test[, knnvector1], train$Life.expectancy, k = 3)
  Knn_Predictions_1 = as.data.frame(predknn1$pred)
  names(Knn_Predictions_1)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_1), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq1 = (Predictions - Actual)^2)

  RMSEknn1 = sqrt(sum((trial$ResidualSq1)/nrow(test)))
  knnrmse1 = c(knnrmse1, RMSEknn1)

  #KNN2
  predknn2 <- knn.reg(train[,knnvector2], test[, knnvector2], train$Life.expectancy, k = 3)
  Knn_Predictions_2 = as.data.frame(predknn2$pred)
  names(Knn_Predictions_2)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_2), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq2 = (Predictions - Actual)^2)

  RMSEknn2 = sqrt(sum((trial$ResidualSq2)/nrow(test)))
  knnrmse2 = c(knnrmse2, RMSEknn2)

  # KNN3
  predknn3 <- knn.reg(train[,knnvector3], test[, knnvector3], train$Life.expectancy, k = 3)
  Knn_Predictions_3 = as.data.frame(predknn3$pred)
  names(Knn_Predictions_3)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_3), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq3 = (Predictions - Actual)^2)

  RMSEknn3 = sqrt(sum((trial$ResidualSq3)/nrow(test)))
  knnrmse3 = c(knnrmse3, RMSEknn3)
  # End For
}

# putting the dataframe together and outputting relevant statistics
Knn.Model.Average.RMSE = cbind(knnrmse1, knnrmse2, knnrmse3)
knnrmsedf = as.data.frame(Knn.Model.Average.RMSE)
Means = colMeans(Knn.Model.Average.RMSE)
SDs = round(colSds(Knn.Model.Average.RMSE), 3)
knnrange1 = max(knnrmsedf$knnrmse1) - min(knnrmsedf$knnrmse1)
knnrange2 = max(knnrmsedf$knnrmse2) - min(knnrmsedf$knnrmse2)
knnrange3 = max(knnrmsedf$knnrmse3) - min(knnrmsedf$knnrmse3)

# Looking at range of each model to determine how variant it is - can also look at stats above
Means
print("Standard Devs")
SDs
print("ranges - measure of variance between train/test shuffles")
print(paste("Model 1 Range" , round(knnrange1, 3)))
print(paste("Model 2 Range" , round(knnrange2, 3)))
print(paste("Model 3 Range" , round(knnrange3, 3)))
cat("Summary: 
    ")
summary(Knn.Model.Average.RMSE)


```

# Regression Tree Model
```{r tree}
#Building a tree out, based on ALL variables from the regression model
par(mfrow=c(1,1))
tree.life<-tree(Life.expectancy~.,train,minsize=5)
summary(tree.life)
plot(tree.life)
text(tree.life,pretty=0)

#Perform CV to deterine if we need to prune the tree. -- 7 parameters was the lowest
set.seed(1234)
cv.life<-cv.tree(tree.life,FUN=prune.tree,method="deviance")
plot(cv.life)
plot(cv.life$size, cv.life$dev, type='b')

#Building a tree out, based on SELECT variables from the regression model
par(mfrow=c(1,1))
tree.life2<-tree(Life.expectancy ~ Schooling + Adult.Mortality + HIV.AIDS + GDP + thinness.5.9.years + Diphtheria + Alcohol,Life14,minsize=5)
summary(tree.life2)
plot(tree.life2)
text(tree.life2,pretty=0)

#Perform CV to deterine if we need to prune the tree. -- All 7 parameters was the lowest
set.seed(1234)
cv.life2<-cv.tree(tree.life2,FUN=prune.tree,method="deviance")
plot(cv.life2)

#Fitting a final model for predicting future values. 
#both versions of the tree identified that 7 parameters was the best fit
prune.life=prune.tree(tree.life,best=7)
plot(prune.life)
text(prune.life,pretty=0)

#Looking at the predictions of the pruned tree
tree.ASE <- mean((test$Life.expectancy - predict(prune.life,newdata =test))^2)
tree.ASE
```
