---
title: "Combined_Stats2_Project1_AC_AP_GL"
author: "Adam Canton, Anish Patel & Grace Lang"
date: "6/7/2020"
output: word_document
---

Kaggle link for data reference: https://www.kaggle.com/kumarajarshi/life-expectancy-who

```{r setup, include=FALSE}

library(class)
library(caret)
library(e1071)
library(magrittr)
library(dplyr)
library(tidyr)
library(naniar)
library(ggplot2)
library(plotly)
library(forcats)
library(ggExtra)
library(glmnet)
library(GGally)
library(ggcorrplot)
library(car)
library(olsrr)
library(rgl)
library(tree)
library(ISLR)
library(leaps)
library(matrixStats)
library(FNN)
library(MASS)
library(reshape2)

#Adam's Connections
life <- read.csv(file = "F:/R For Real/Stats2 Project 1/Life Expectancy Data.csv")

#Anish's Connections
#life <- read.csv("Life Expectancy Data.csv")

#Grace's Connections
#life <-read.csv("C:/Users/david/OneDrive/Desktop/AppliedStats/6372_AppliedStats_GraceLang/Project1/LifeExpectancyData.csv")
#life <-read.csv("C:/Users/Dave/Desktop/AppliedStats/Stats2_Project1/LifeExpectancyData.csv")


#Just pull out 2014 into the data
Life14 <- subset(life, Year==2014)
summary(Life14)
```
#Column definitions:
Adult.mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)
Infant.death: Number of Infant Deaths per 1000 population
Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)
percentage expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita(%)
Hepatitis B: Hepatitis B (HepB) immunization coverage among 1-year-olds (%)
Measles - number of reported cases per 1000 population
BMI: Average Body Mass Index of entire population
Under-five-deaths: Number of under-five deaths per 1000 population
Polio: Polio (Pol3) immunization coverage among 1-year-olds (%)
Total expenditure: General government expenditure on health as a percentage of total government expenditure (%)
Diphtheria: Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds (%)
HIV/AIDS: Deaths per 1 000 live births HIV/AIDS (0-4 years)
GDP: Gross Domestic Product per capita (in USD)
Population: Population of the country
thinness: Prevalence of thinness among children and adolescents for Age 10 to 19 (% )
Thinness 5-9: Prevalence of thinness among children for Age 5 to 9(%)
income composition: Human Development Index in terms of income composition of resources (index ranging from 0 to 1)
Schooling: Number of years of Schooling(years)


##Things to note from summary stats:
 * 83% of data comes from developing countries, life exp may be lower.
 * Mean life expectancy is 71.54 yr old
 * The max for Measles seems outrageously high 79k - which country? outlier?
 * Looks like there's a couple of positive correlations with life expectancy: income.compostion & schooling
 * Adult mortality seems to have a gender split, but cannot determine for certain
 * No outward evidence of any of the parameters having a non-linear relationship (adding a quadratic)
 
 # Fixing NA's in GDP
```{r}
Life14$GDP[159] = 2520.2027
Life14$GDP[11] = 29436.13
Life14$GDP[20] = 3081.23
Life14$GDP[28] = 1559.38
Life14$GDP[39] = 6208.58
Life14$GDP[44] = 19771.65
Life14$GDP[45] = 1568.63
Life14$GDP[46] = 352.82
Life14$GDP[51] = 3378.68
Life14$GDP[54] = 1209.68
Life14$GDP[61] = 607.21
Life14$GDP[77] = 5608.62
Life14$GDP[89] = 1279.64
Life14$GDP[90] = 1998.49
Life14$GDP[95] = 6466.52
Life14$GDP[107] = 2971.73
Life14$GDP[131] = 27802.96
Life14$GDP[132] = 2674.35
Life14$GDP[136] = 8743.89
Life14$GDP[137] = 6684.67
Life14$GDP[146] = 18671.59
Life14$GDP[162] = 5487.92
Life14$GDP[173] = 47614.61
Life14$GDP[174] = 1000.00
Life14$GDP[175] = 55025.13
Life14$GDP[179] = 16053.24
Life14$GDP[180] = 2030.31
Life14$GDP[181] = 1674.28
```
# Fixing Population NA's
```{r}
Life14$Population[159] = 18715672
Life14$Population[11] = 370633
Life14$Population[20] = 10710000
Life14$Population[28] = 22650000
Life14$Population[39] = 5244359
Life14$Population[44] = 10510000
Life14$Population[45] = 25500000
Life14$Population[46] = 10780263
Life14$Population[51] = 90420000
Life14$Population[54] = 5580000
Life14$Population[61] = 2024000
Life14$Population[77] = 77470000
Life14$Population[89] = 5836000
Life14$Population[90] = 6640000
Life14$Population[95] = 6362000
Life14$Population[107] = 107446
Life14$Population[131] = 50750000
Life14$Population[132] = 3556000
Life14$Population[136] = 178296
Life14$Population[137] = 108861
Life14$Population[146] = 5420000
Life14$Population[162] = 2070000
Life14$Population[173] = 64350000
Life14$Population[174] = 49960000
Life14$Population[175] = 318400000
Life14$Population[179] = 30050000
Life14$Population[180] = 91710000
Life14$Population[181] = 25820000
Life14$Population[130] = 2459000
Life14$Population[145] = 5470000
Life14$Population[172] = 9214000
Life14$Population[88] = 3691000
Life14$Population[12] = 1336000
Life14$Population[140] = 30920000
Life14$Population[14] = 284825
Life14$Population[5] = 92562
Life14$Population[66] = 108902
Life14$Population[116] = 4510000
Life14$Population[24] = 409769
Life14$Population[121] = 4027000
Life14$Population[149] = 13420000
Life14$Population[42] = 11310000
```
# Fixing 27 - 0 Percent Expenditures
```{r}
Life14$percentage.expenditure[11] = 748.36
Life14$percentage.expenditure[20] = 115.35
Life14$percentage.expenditure[28] = 16.85
Life14$percentage.expenditure[39] = 34.78
Life14$percentage.expenditure[44] = 1249.23
Life14$percentage.expenditure[46] = 2.68
Life14$percentage.expenditure[51] = 48.66
Life14$percentage.expenditure[54] = 2.82
Life14$percentage.expenditure[61] = 4.93
Life14$percentage.expenditure[77] = 192.78
Life14$percentage.expenditure[89] = 35.5
Life14$percentage.expenditure[90] = 13.79
Life14$percentage.expenditure[95] = 197.84
Life14$percentage.expenditure[107] = 102.43
Life14$percentage.expenditure[131] = 1093.43
Life14$percentage.expenditure[132] = 113.87
Life14$percentage.expenditure[136] = 189.84
Life14$percentage.expenditure[137] = 178.51
Life14$percentage.expenditure[146] = 1024.94
Life14$percentage.expenditure[159] = 31.65
Life14$percentage.expenditure[162] = 219.74
Life14$percentage.expenditure[173] = 3682.48
Life14$percentage.expenditure[174] = 11.47
Life14$percentage.expenditure[175] = 4541.8
Life14$percentage.expenditure[179] = 235.35
Life14$percentage.expenditure[180] = 48.82
Life14$percentage.expenditure[181] = 12.9
```
# Fixing the 10 missing school variables 
```{r}
Life14$Schooling[28] = 5.2
Life14$Schooling[44] = 12.7
Life14$Schooling[45] = 10.8
Life14$Schooling[46] = 6.8
Life14$Schooling[131] = 12.2
Life14$Schooling[132] = 11.6
Life14$Schooling[149] = 0
Life14$Schooling[173] = 17.4
Life14$Schooling[174] = 8
Life14$Schooling[175] = 16.3
```
#What variables are still missing
```{r}
gg_miss_var(Life14)
```

```{r}
#Can I add the avgs into the NAs without a lot of shift in data
#Life14 %>% ggplot(aes(x = Income.composition.of.resources)) + geom_histogram() #normally distributed, low effect 2% change in mean
#Life14 %>% ggplot(aes(x = Hepatitis.B)) + geom_histogram() #left-skewed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = Total.expenditure)) + geom_histogram() #normally distributed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = thinness.5.9.years)) + geom_histogram() #right-skewerd, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = thinness..1.19.years)) + geom_histogram() #right-skewerd, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = BMI)) + geom_histogram() #normally distributed, low effect 0% change in mean
#Life14 %>% ggplot(aes(x = Alcohol)) + geom_histogram() #right-skewerd, low effect 0% change in mean

#Replacing NAs with the mean of the column
Life14$Income.composition.of.resources[is.na(Life14$Income.composition.of.resources)] <- round(mean(Life14$Income.composition.of.resources, na.rm = TRUE))
Life14$Hepatitis.B[is.na(Life14$Hepatitis.B)] <- round(mean(Life14$Hepatitis.B, na.rm = TRUE))
Life14$Total.expenditure[is.na(Life14$Total.expenditure)] <- round(mean(Life14$Total.expenditure, na.rm = TRUE))
Life14$thinness..1.19.years[is.na(Life14$thinness..1.19.years)] <- round(mean(Life14$thinness..1.19.years, na.rm = TRUE))
Life14$thinness.5.9.years[is.na(Life14$thinness.5.9.years)] <- round(mean(Life14$thinness.5.9.years, na.rm = TRUE))
Life14$BMI[is.na(Life14$BMI)] <- round(mean(Life14$BMI, na.rm = TRUE))
Life14$Alcohol[is.na(Life14$Alcohol)] <- round(mean(Life14$Alcohol, na.rm = TRUE))


```
# Adding in some feature engineering
```{r}
#Factorizing things
IncomeCompFactor = cut(Life14$Income.composition.of.resources, breaks = c(0,0.4,0.7,1), labels = c("Low", "Moderate", "High"))

HepBfactor = cut(Life14$Hepatitis.B, breaks = c(0,25,50,75,100), labels = c("Deficient", "Low", "Moderate", "High"))

PercentExpendFactor = cut(Life14$percentage.expenditure, breaks = c(0,5000,15000,20000), labels = c("Low", "Moderate", "High"))

PopulationFactor = cut(Life14$Population, breaks = c(0, 3.612e+05,2.196e+07,1.294e+09), labels = c("Low", "Moderate", "High"))

GdpFactor = cut(Life14$GDP, breaks = c(0,40000,80000,120000), labels = c("Low","Moderate","High"))

SchoolFactor = cut(Life14$Schooling, breaks = c(0,12,16,21), labels = c("Low","Moderate", "High"))

# Binding them back to Life14
Life14 = cbind(Life14, HepBfactor, PercentExpendFactor, PopulationFactor, GdpFactor, SchoolFactor, IncomeCompFactor)

#Fixing some errors where for some reason it wasn't picking up these 3
Life14$PercentExpendFactor[45] = as.factor("Low")
Life14$PercentExpendFactor[149] = as.factor("Low")
Life14$SchoolFactor[149] = as.factor("Low")

#adding log variables
Life14 = Life14 %>% mutate(LHIV.AIDS = log(HIV.AIDS))
Life14 = Life14 %>% mutate(Linfant.deaths = log(infant.deaths + 1))
Life14 = Life14 %>% mutate(LUnder5Deaths = log(under.five.deaths + 1))
Life14 = Life14 %>% mutate(LThin19 = log(thinness..1.19.years + 1))
Life14 = Life14 %>% mutate(LGDP = log(GDP))

```


#Visualizing some of the data
```{r}
# Life expectancy by Country and status -  goes well with boxplot below 
Life14 %>% group_by(Country, Status) %>% summarise(LifeEx = Life.expectancy) %>%
  ggplot(aes(x = reorder(Country, -LifeEx), y = LifeEx, fill = Status)) +
  geom_col(width = 0.75, na.rm = TRUE) +
  theme(axis.text.x = element_blank()) + 
  xlab("Country") + 
  ggtitle("Life Expectancy 2014")
```

Boxplot of life expectancy by country status:
```{r}
Life14 %>% ggplot(aes(y = Life.expectancy)) + geom_boxplot() + facet_wrap(~Status)
```

Pairwise plots for selected variables:
```{r message = FALSE, warning = FALSE}
df2 <- dplyr::select(Life14, c(Life.expectancy, Adult.Mortality, Total.expenditure, HIV.AIDS, Income.composition.of.resources, Schooling))
library(GGally)
ggpairs(df2)
```

Correlation between life expectancy and all the numeric independent variables:
```{r}
# Error in Cor.xy - x must be numeric
cor.xy <- cor(Life14 %>% dplyr::select(-c(Country, Year, Status,IncomeCompFactor,HepBfactor,PercentExpendFactor,PopulationFactor,GdpFactor,SchoolFactor)), use = "complete.obs")
LE.cor <- data.frame(var = rownames(cor.xy)[-1], correlation = cor.xy[-1, 1])
LE.cor %>% ggplot(aes(x = var, y = correlation)) + geom_col() + coord_flip()
```


# Checking Correlations between original variables and their logged counterparts as well as the response - we may only achieve some minor inprovements in GDP and HIV.AIDS but under 5 deaths and infant deaths have changed significantly. Also HIV.AIDS and infant deaths without the log are oddly related
```{r}
corr <- Life14 %>% dplyr::select(Life.expectancy, LHIV.AIDS, Linfant.deaths, LUnder5Deaths,
                          LThin19, LGDP, GDP, infant.deaths, HIV.AIDS, under.five.deaths, thinness..1.19.years) 

corr <- round(cor(corr), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE, lab_size = 3, method = "circle",
           colors = c("tomato2", "white", "springgreen3"),
           title = "Correlations of all relevant variables",
           ggtheme = theme_bw())
```


# Looking at Correlations for the rest of the predictors
```{r}
exclude_factors <- c( "Country", "Year", "Status", "thinness..1.19.years", "under.five.deaths", "IncomeCompFactor",
                     "HepBfactor", "PercentExpendFactor", "SchoolFactor", "GdpFactor", "PopulationFactor",
                     "LHIV.AIDS", "Linfant.deaths", "LUnder5Deaths", "LThin19", "LGDP")

corr <- Life14 %>% dplyr::select(-all_of(exclude_factors))

corr <- round(cor(corr), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE, lab_size = 3, method = "circle",
           colors = c("tomato2", "white", "springgreen3"),
           title = "Correlations of all relevant variables",
           ggtheme = theme_bw())
```

Get rid of Country Name and Year
```{r}
exclude_factors <- c("Country", "Year")
Life14 <- Life14 %>% dplyr::select( -all_of(exclude_factors))
```


```{r}
# Set training and Testing Sets
set.seed(1234)
index<-sample(1:dim(Life14)[1],90,replace=F)
train<-Life14[index,]
test<-Life14[-index,]

# Forward Selection 
reg.fwd=regsubsets(Life.expectancy~.,data=train,method="forward",nvmax=20)
bics<-summary(reg.fwd)$bic
plot(1:21,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

# Adjr2
adjr2<-summary(reg.fwd)$adjr2
plot(1:21,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)


MallowCP <- summary(reg.fwd)$cp
plot(1:21,MallowCP,type="l",ylab="Mallow's CP",xlab="# of predictors")
index<-which(MallowCP==min(MallowCP))
points(index,MallowCP[index],col="red",pch=10)
```



# Used to quickly cycle though models while shuffling train/test sets
Ols step both does forward and backward selection at each step in this case based on model AIC
```{r, warning=FALSE}
index<-sample(1:dim(Life14)[1],128,replace=F)
train<-Life14[index,]
test<-Life14[-index,]

modelgenerator <- lm(Life.expectancy ~ . , data = train)
ols_step_both_aic(modelgenerator)

# Gets significant interactions based on aic - can take a minute to run
#Interaction.modelgenerator <- lm(Life.expectancy ~ (.)^2 , data = train)
#ols_step_both_aic(Interaction.modelgenerator)
```


# Gut Model - EDA first - human variable selection - Some tech use to look at optimum number of predictors and comparison of models - income comp increases model variance
```{r}

modeltest1 <- lm(Life.expectancy ~ SchoolFactor + LHIV.AIDS + Adult.Mortality + Linfant.deaths , data = train)


summary(modeltest1)
vif(modeltest1)
ols_plot_resid_fit(modeltest1)
ols_plot_resid_lev(modeltest1)
ols_plot_resid_qq(modeltest1)
ols_plot_resid_hist(modeltest1)
ols_plot_cooksd_bar(modeltest1)
```


# complex model - Contains 4 quadratic terms and 1 interaction. Adding more interactions seems to overfit. Will occasionally show some crazy Cook's D based on
which train/test gets selected. For a better view - run the model benchmark - easy to see over 100-500 interations
```{r}
modeltest2 <-  lm(Life.expectancy ~  SchoolFactor + poly(LHIV.AIDS,2) + poly(Adult.Mortality,2) + poly(under.five.deaths,2) + poly(thinness..1.19.years,2) +
                   LGDP + Income.composition.of.resources + percentage.expenditure + Income.composition.of.resources:Schooling, data = train)
summary(modeltest2)
ols_plot_resid_fit(modeltest2)
ols_plot_resid_lev(modeltest2)
ols_plot_resid_qq(modeltest2)
ols_plot_resid_hist(modeltest2)
ols_plot_cooksd_bar(modeltest2)
```


# Linear Model Benchmarking - Welch's t-test run at the end (likely a conservative option). 500 observations each (Robust to normal violations) - however sample SD can vary a decent bit based on train/test shuffles and the whole population is 128c55....
```{r, warning=FALSE}
# Set number of times you would like to repeat the sampling/testing 
iterations = 1:500

# the initial values for the columns (might not need these now that ive switched to building columns) 
rmseSimple = c()
rmseComplex = c()


# Start of Loop
for(i in iterations){
  # Resets sample every iteration 
  index<- sample(1:dim(Life14)[1],128,replace=F)
  train<- Life14[index,]
  test<- Life14[-index,]
  

  # the model runs 
  modeltest1 
  
  modeltest2 
  
  
  # predictors and column building
  predictions1 <- modeltest1 %>% predict(test)
  
  d1 = data.frame(R2 = R2(predictions1,test$Life.expectancy),
                  RMSE = RMSE(predictions1,test$Life.expectancy), MAE = MAE(predictions1, test$Life.expectancy))
  rmseSimple = c(rmseSimple,d1$RMSE)
  
  predictions2 <- modeltest2 %>% predict(test)

  d2 = data.frame(R2 = R2(predictions2,test$Life.expectancy),
                RMSE = RMSE(predictions2,test$Life.expectancy), MAE = MAE(predictions2, test$Life.expectancy))
  rmseComplex = c(rmseComplex, d2$RMSE)
  
 
  # End for
}
# putting the dataframe together and outputting relevant statistics
Model.Average.RMSE = cbind(rmseSimple, rmseComplex)
rmsedf = as.data.frame(Model.Average.RMSE)
Means = colMeans(Model.Average.RMSE)
SDs = round(colSds(Model.Average.RMSE), 3)
range1 = max(rmsedf$rmseSimple) - min(rmsedf$rmseSimple)
range2 = max(rmsedf$rmseComplex) - min(rmsedf$rmseComplex)
rmsedf1 = melt(rmsedf,rmse = c("n", "rmse"))

# Looking at descriptive stats
Means
print("Standard Devs")
SDs
print("ranges - measure of variance between train/test shuffles")
print(paste("Model 1 Range" , round(range1, 3)))
print(paste("Model 2 Range" , round(range2, 3)))
cat("Summary: 
    ")
summary(Model.Average.RMSE)


# Column
rmsedf1 %>% group_by(variable) %>% summarise(mean = (mean(value))) %>% 
  ggplot(aes(x = reorder(variable, -mean), y = mean, fill = variable)) + geom_col(width = 0.75) + geom_text(aes(label = round(mean,3), vjust = -0.5)) + 
  ggtitle("Average RMSE over 100 Shuffles (Linear Models)") + xlab("Model #") + ylab("Mean RMSE")

# Boxplot
rmsedf1 %>%  ggplot(aes(x = variable, y = value)) + geom_boxplot() + facet_wrap(~variable,ncol = TRUE) +
  ggtitle("Mean RMSE Distribution by Model") + ylab("Mean RMSE") + coord_flip() + 
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())

# Here we can see there is no significant difference between the models in terms of RMSE
t.test(rmseSimple,rmseComplex, var.equal = FALSE)
```

```{r}
# Creating scaled variables for knn testing
Life14Scale = Life14
Life14Scale$Life.expectancy = scale(Life14Scale$Life.expectancy) 
Life14Scale$Adult.Mortality = scale(Life14Scale$Adult.Mortality)
Life14Scale$percentage.expenditure = scale(Life14Scale$percentage.expenditure)
Life14Scale$Diphtheria = scale(Life14Scale$Diphtheria)
Life14Scale$Hepatitis.B = scale(Life14Scale$Hepatitis.B)
Life14Scale$Income.composition.of.resources = scale(Life14Scale$Income.composition.of.resources)
Life14Scale$Schooling = scale(Life14Scale$Schooling)
Life14Scale$LHIV.AIDS = scale(Life14Scale$LHIV.AIDS)
Life14Scale$infant.deaths = scale(Life14Scale$infant.deaths)
Life14Scale$Linfant.deaths = scale(Life14Scale$Linfant.deaths)
Life14Scale$LUnder5Deaths = scale(Life14Scale$LUnder5Deaths)
Life14Scale$LThin19 = scale(Life14Scale$LThin19)
Life14Scale$LGDP = scale(Life14Scale$LGDP)
Life14Scale$Total.expenditure = scale(Life14Scale$Total.expenditure)
```


Knn Model benchmarking
```{r, warning=FALSE}
# Set knn vectors - these are the column #s that will be analyzed 
knnvector1 = c(3,4,14,20) # Adult.Mortality, infant, hiv, schooling
knnvector2 = c(3,20,27,28) # Adult.Mortality, Schooling, Log(HIV), Log(infant)
knnvector3 = c(3,6,13,20) # Adult.Mortality, percentage.expenditure, Diptheria, schooling

# Start of Loop Variables
iterations = 1:500
i = 0
knnrmse1 = c()
knnrmse2 = c()
knnrmse3 = c()

# Loop start
for(i in iterations){
  # Create train and test sets
  index<-sample(1:dim(Life14Scale)[1],128,replace=F)
  train<-Life14Scale[index,]
  test<-Life14Scale[-index,]


  #KNN1
  predknn1 <- knn.reg(train[,knnvector1], test[, knnvector1], train$Life.expectancy, k = 4)
  Knn_Predictions_1 = as.data.frame(predknn1$pred)
  names(Knn_Predictions_1)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_1), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq1 = (Predictions - Actual)^2)

  RMSEknn1 = sqrt(sum((trial$ResidualSq1)/nrow(test)))
  knnrmse1 = c(knnrmse1, RMSEknn1)

  #KNN2
  predknn2 <- knn.reg(train[,knnvector2], test[, knnvector2], train$Life.expectancy, k = 9)
  Knn_Predictions_2 = as.data.frame(predknn2$pred)
  names(Knn_Predictions_2)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_2), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq2 = (Predictions - Actual)^2)

  RMSEknn2 = sqrt(sum((trial$ResidualSq2)/nrow(test)))
  knnrmse2 = c(knnrmse2, RMSEknn2)

  # KNN3
  predknn3 <- knn.reg(train[,knnvector3], test[, knnvector3], train$Life.expectancy, k = 4)
  Knn_Predictions_3 = as.data.frame(predknn3$pred)
  names(Knn_Predictions_3)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions_3), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq3 = (Predictions - Actual)^2)

  RMSEknn3 = sqrt(sum((trial$ResidualSq3)/nrow(test)))
  knnrmse3 = c(knnrmse3, RMSEknn3)
  # End For
}

# putting the dataframe together and outputting relevant statistics
Knn.Model.Average.RMSE = cbind(knnrmse1, knnrmse2, knnrmse3)
knnrmsedf = as.data.frame(Knn.Model.Average.RMSE)
Means = colMeans(Knn.Model.Average.RMSE)
SDs = round(colSds(Knn.Model.Average.RMSE), 3)
knnrange1 = max(knnrmsedf$knnrmse1) - min(knnrmsedf$knnrmse1)
knnrange2 = max(knnrmsedf$knnrmse2) - min(knnrmsedf$knnrmse2)
knnrange3 = max(knnrmsedf$knnrmse3) - min(knnrmsedf$knnrmse3)
knnrmsedf1 = melt(knnrmsedf,rmse = c("n", "rmse"))

# Looking at descriptive stats
Means
print("Standard Devs")
SDs
print("ranges - measure of variance between train/test shuffles")
print(paste("Model 1 Range" , round(knnrange1, 4)))
print(paste("Model 2 Range" , round(knnrange2, 4)))
print(paste("Model 3 Range" , round(knnrange3, 4)))
cat("Summary: 
    ")
summary(Knn.Model.Average.RMSE)


# Column
knnrmsedf1 %>% group_by(variable) %>% summarise(mean = (mean(value))) %>% 
  ggplot(aes(x = reorder(variable, -mean), y = mean, fill = variable)) + geom_col(width = 0.75) + geom_text(aes(label = round(mean,3), vjust = -0.5)) + 
  ggtitle("Average RMSE over 100 Shuffles (KNN)") + xlab("Model #") + ylab("Mean RMSE")

# Boxplot
knnrmsedf1 %>%  ggplot(aes(x = variable, y = value)) + geom_boxplot() + facet_wrap(~variable,ncol = TRUE) +
  ggtitle("Mean RMSE Distribution by Model") + ylab("Mean RMSE") + coord_flip() + 
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())

# 
# Dont forget multiple comparison adjustment

t.test(knnrmse1,knnrmse2, var.equal = FALSE)
t.test(knnrmse1,knnrmse3, var.equal = FALSE)
t.test(knnrmse2,knnrmse3, var.equal = FALSE)
```


# Looks for optimal K value between 1-50. Just change the knnvector to run different models.
```{r}
iterations = 50
numks = 50
masterRMSE = matrix(nrow = iterations, ncol = numks)


for(j in 1:iterations)
{
rmseknn = data.frame(rmse = numeric(50), k = numeric(50))
index<-sample(1:dim(Life14Scale)[1],128,replace=F)
train<-Life14Scale[index,]
test<-Life14Scale[-index,]
for(i in 3:numks)
{
  predknn <- knn.reg(train[,knnvector3], test[,knnvector3], train$Life.expectancy, k = i)
  Knn_Predictions = as.data.frame(predknn$pred)
  names(Knn_Predictions)[1] <- "Predictions"

  trial <- data.frame(Predictions = c(Knn_Predictions), Actual = c(test$Life.expectancy))

  trial = trial %>% mutate(ResidualSq = (Predictions - Actual)^2)

  RMSEknn = sqrt(sum((trial$ResidualSq)/nrow(test)))
  masterRMSE[j,i] = RMSEknn
}
}
masterRMSE = masterRMSE[,-(1:2)]
MeanRMSE = colMeans(masterRMSE)
# How can I title this plot and change the x axis label? I Suck at Base R
plot(seq(3,numks,1),MeanRMSE, type = "l")
which.min(MeanRMSE)
min(MeanRMSE)
```

This model is here to run against the knn for comparison. It uses the same scaled variables as the knnvector3 (linear version of the KNN).
```{r}
#Create train and test sets for scaled linear models - currently reflects knnvector3
indexS<-sample(1:dim(Life14Scale)[1],128,replace=F)
trainS<-Life14Scale[indexS,]
testS<-Life14Scale[-indexS,]
  
modeltest3 <- lm(Life.expectancy ~ Adult.Mortality +  percentage.expenditure + Diphtheria + Schooling, data = trainS)

summary(modeltest3)
vif(modeltest3)
ols_plot_resid_fit(modeltest3)
ols_plot_resid_lev(modeltest3)
ols_plot_resid_qq(modeltest3)
ols_plot_resid_hist(modeltest3)
ols_plot_cooksd_bar(modeltest3)
```

# Regression Tree Model
```{r tree}
#Building a tree out, based on ALL variables from the regression model
par(mfrow=c(1,1))
tree.life<-tree(Life.expectancy~.,train,minsize=5)
summary(tree.life)
plot(tree.life)
text(tree.life,pretty=0)

#Perform CV to deterine if we need to prune the tree. -- 7 parameters was the lowest
set.seed(1234)
cv.life<-cv.tree(tree.life,FUN=prune.tree,method="deviance")
plot(cv.life)
plot(cv.life$size, cv.life$dev, type='b')

#Building a tree out, based on SELECT variables from the regression model
par(mfrow=c(1,1))
tree.life2<-tree(Life.expectancy ~ Schooling + Adult.Mortality + HIV.AIDS + GDP + thinness.5.9.years + Diphtheria + Alcohol,Life14,minsize=5)
summary(tree.life2)
plot(tree.life2)
text(tree.life2,pretty=0)

#Perform CV to deterine if we need to prune the tree. -- All 7 parameters was the lowest
set.seed(1234)
cv.life2<-cv.tree(tree.life2,FUN=prune.tree,method="deviance")
plot(cv.life2)

#Fitting a final model for predicting future values. 
#both versions of the tree identified that 7 parameters was the best fit
prune.life=prune.tree(tree.life,best=7)
plot(prune.life)
text(prune.life,pretty=0)

#Looking at the predictions of the pruned tree
tree.ASE <- mean((test$Life.expectancy - predict(prune.life,newdata =test))^2)
tree.ASE
```
