---
title: "Project1_AppliedStats_GraceLang"
author: "Grace Lang"
date: "5/30/2020"
output: word_document
---

---
title: "Project1_GraceLang"
author: "Grace Lang"
date: "5/24/2020"
output: html_document
---

Kaggle link for data reference: https://www.kaggle.com/kumarajarshi/life-expectancy-who


```{r setup, include=FALSE}
#life <-read.csv("C:/Users/david/OneDrive/Desktop/AppliedStats/6372_AppliedStats_GraceLang/Project1/LifeExpectancyData.csv")
life <-read.csv("C:/Users/Dave/Desktop/AppliedStats/Stats2_Project1/LifeExpectancyData.csv")

#Just pull out 2014 into the data
Life14 <- subset(life, Year==2014)
summary(Life14)
```
#Column definitions:
Adult.mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)
Infant.death: Number of Infant Deaths per 1000 population
Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)
percentage expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita(%)
Hepatitis B: Hepatitis B (HepB) immunization coverage among 1-year-olds (%)
Measles - number of reported cases per 1000 population


Things to note from summary stats:
 * 83% of data comes from developing countries, life exp may be lower.
 * Mean life expectancy is 71.54 yr old
 * The max for Measles seems outrageously high 79k - which country? outlier?
 * Looks like there's a couple of positive correlations with life expectancy: income.compostion & schooling
```{r plots}
#Review the correlaton of data to life expectancy
par(mfrow=c(3,6))
plot(Life14$Life.expectancy,Life14$Adult.Mortality, xlab="Life Expect",ylab="Adult Mortality")
plot(Life14$Life.expectancy,Life14$infant.deaths, xlab="Life Expect",ylab="Infant death")
plot(Life14$Life.expectancy,Life14$Alcohol , xlab="Life Expect",ylab="Alcohol")
plot(Life14$Life.expectancy,Life14$percentage.expenditure , xlab="Life Expect",ylab="percentage.expenditure")
plot(Life14$Life.expectancy,Life14$Hepatitis.B , xlab="Life Expect",ylab="Hepatitis.B")
plot(Life14$Life.expectancy,Life14$Measles , xlab="Life Expect",ylab="Measles")
plot(Life14$Life.expectancy,Life14$BMI , xlab="Life Expect",ylab="BMI")
plot(Life14$Life.expectancy,Life14$under.five.deaths , xlab="Life Expect",ylab="under.five.deaths")
plot(Life14$Life.expectancy,Life14$Polio , xlab="Life Expect",ylab="Polio")
plot(Life14$Life.expectancy,Life14$Total.expenditure , xlab="Life Expect",ylab="Total.expenditure")
plot(Life14$Life.expectancy,Life14$Diphtheria , xlab="Life Expect",ylab="Diphtheria")
plot(Life14$Life.expectancy,Life14$HIV.AIDS , xlab="Life Expect",ylab="HIV.AIDS")
plot(Life14$Life.expectancy,Life14$GDP , xlab="Life Expect",ylab="GDP")
plot(Life14$Life.expectancy,Life14$Population , xlab="Life Expect",ylab="Population")
plot(Life14$Life.expectancy,Life14$thinness..1.19.years , xlab="Life Expect",ylab="thinness..1.19.years")
plot(Life14$Life.expectancy,Life14$thinness.5.9.years , xlab="Life Expect",ylab="thinness.5.9.years")
plot(Life14$Life.expectancy,Life14$Income.composition.of.resources , xlab="Life Expect",ylab="Income.composition.of.resources")
plot(Life14$Life.expectancy,Life14$Schooling , xlab="Life Expect",ylab="Schooling")
```

```{r linear}
lm.life <- lm(Life.expectancy~.,data=Life14)
plot(lm.life$residuals)
```