---
title: "Project1_AppliedStats_GraceLang"
author: "Grace Lang"
date: "5/30/2020"
output: word_document
---

Kaggle link for data reference: https://www.kaggle.com/kumarajarshi/life-expectancy-who


```{r setup, include=FALSE}
life <-read.csv("C:/Users/david/OneDrive/Desktop/AppliedStats/6372_AppliedStats_GraceLang/Project1/LifeExpectancyData.csv")
#life <-read.csv("C:/Users/Dave/Desktop/AppliedStats/Stats2_Project1/LifeExpectancyData.csv")

#Just pull out 2014 into the data
Life14 <- subset(life, Year==2014)
summary(Life14)
```
#Column definitions:
Adult.mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)
Infant.death: Number of Infant Deaths per 1000 population
Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)
percentage expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita(%)
Hepatitis B: Hepatitis B (HepB) immunization coverage among 1-year-olds (%)
Measles - number of reported cases per 1000 population
BMI: Average Body Mass Index of entire population
Under-five-deaths: Number of under-five deaths per 1000 population
Polio: Polio (Pol3) immunization coverage among 1-year-olds (%)
Total expenditure: General government expenditure on health as a percentage of total government expenditure (%)
Diphtheria: Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds (%)
HIV/AIDS: Deaths per 1 000 live births HIV/AIDS (0-4 years)
GDP: Gross Domestic Product per capita (in USD)
Population: Population of the country
thinness: Prevalence of thinness among children and adolescents for Age 10 to 19 (% )
Thinness 5-9: Prevalence of thinness among children for Age 5 to 9(%)
income composition: Human Development Index in terms of income composition of resources (index ranging from 0 to 1)
Schooling: Number of years of Schooling(years)


##Things to note from summary stats:
 * 83% of data comes from developing countries, life exp may be lower.
 * Mean life expectancy is 71.54 yr old
 * The max for Measles seems outrageously high 79k - which country? outlier?
 * Looks like there's a couple of positive correlations with life expectancy: income.compostion & schooling
 * Adult mortality seems to have a gender split, but cannot determine for certain
 * No outward evidence of any of the parameters having a non-linear relationship (adding a quadratic)
```{r plots}
#Review the correlaton of data to life expectancy
par(mfrow=c(3,6))
plot(Life14$Life.expectancy,Life14$Adult.Mortality, xlab="Life Expect",ylab="Adult Mortality")
plot(Life14$Life.expectancy,Life14$infant.deaths, xlab="Life Expect",ylab="Infant death")
plot(Life14$Life.expectancy,Life14$Alcohol , xlab="Life Expect",ylab="Alcohol")
plot(Life14$Life.expectancy,Life14$percentage.expenditure , xlab="Life Expect",ylab="percentage.expenditure")
plot(Life14$Life.expectancy,Life14$Hepatitis.B , xlab="Life Expect",ylab="Hepatitis.B")
plot(Life14$Life.expectancy,Life14$Measles , xlab="Life Expect",ylab="Measles")
plot(Life14$Life.expectancy,Life14$BMI , xlab="Life Expect",ylab="BMI")
plot(Life14$Life.expectancy,Life14$under.five.deaths , xlab="Life Expect",ylab="under.five.deaths")
plot(Life14$Life.expectancy,Life14$Polio , xlab="Life Expect",ylab="Polio")
plot(Life14$Life.expectancy,Life14$Total.expenditure , xlab="Life Expect",ylab="Total.expenditure")
plot(Life14$Life.expectancy,Life14$Diphtheria , xlab="Life Expect",ylab="Diphtheria")
plot(Life14$Life.expectancy,Life14$HIV.AIDS , xlab="Life Expect",ylab="HIV.AIDS")
plot(Life14$Life.expectancy,Life14$GDP , xlab="Life Expect",ylab="GDP")
plot(Life14$Life.expectancy,Life14$Population , xlab="Life Expect",ylab="Population")
plot(Life14$Life.expectancy,Life14$thinness..1.19.years , xlab="Life Expect",ylab="thinness..1.19.years")
plot(Life14$Life.expectancy,Life14$thinness.5.9.years , xlab="Life Expect",ylab="thinness.5.9.years")
plot(Life14$Life.expectancy,Life14$Income.composition.of.resources , xlab="Life Expect",ylab="Income.composition.of.resources")
plot(Life14$Life.expectancy,Life14$Schooling , xlab="Life Expect",ylab="Schooling")

```

# Fixing NA's in GDP
```{r}
Life14$GDP[159] = 2520.2027
Life14$GDP[11] = 29436.13
Life14$GDP[20] = 3081.23
Life14$GDP[28] = 1559.38
Life14$GDP[39] = 6208.58
Life14$GDP[44] = 19771.65
Life14$GDP[45] = 1568.63
Life14$GDP[46] = 352.82
Life14$GDP[51] = 3378.68
Life14$GDP[54] = 1209.68
Life14$GDP[61] = 607.21
Life14$GDP[77] = 5608.62
Life14$GDP[89] = 1279.64
Life14$GDP[90] = 1998.49
Life14$GDP[95] = 6466.52
Life14$GDP[107] = 2971.73
Life14$GDP[131] = 27802.96
Life14$GDP[132] = 2674.35
Life14$GDP[136] = 8743.89
Life14$GDP[137] = 6684.67
Life14$GDP[146] = 18671.59
Life14$GDP[162] = 5487.92
Life14$GDP[173] = 47614.61
Life14$GDP[174] = 1000.00
Life14$GDP[175] = 55025.13
Life14$GDP[179] = 16053.24
Life14$GDP[180] = 2030.31
Life14$GDP[181] = 1674.28
```

```{r}
# Fixing Population NA's
Life14$Population[159] = 18715672
Life14$Population[11] = 370633
Life14$Population[20] = 10710000
Life14$Population[28] = 22650000
Life14$Population[39] = 5244359
Life14$Population[44] = 10510000
Life14$Population[45] = 25500000
Life14$Population[46] = 10780263
Life14$Population[51] = 90420000
Life14$Population[54] = 5580000
Life14$Population[61] = 2024000
Life14$Population[77] = 77470000
Life14$Population[89] = 5836000
Life14$Population[90] = 6640000
Life14$Population[95] = 6362000
Life14$Population[107] = 107446
Life14$Population[131] = 50750000
Life14$Population[132] = 3556000
Life14$Population[136] = 178296
Life14$Population[137] = 108861
Life14$Population[146] = 5420000
Life14$Population[162] = 2070000
Life14$Population[173] = 64350000
Life14$Population[174] = 49960000
Life14$Population[175] = 318400000
Life14$Population[179] = 30050000
Life14$Population[180] = 91710000
Life14$Population[181] = 25820000
Life14$Population[130] = 2459000
Life14$Population[145] = 5470000
Life14$Population[172] = 9214000
Life14$Population[88] = 3691000
Life14$Population[12] = 1336000
Life14$Population[140] = 30920000
Life14$Population[14] = 284825
Life14$Population[5] = 92562
Life14$Population[66] = 108902
Life14$Population[116] = 4510000
Life14$Population[24] = 409769
Life14$Population[121] = 4027000
Life14$Population[149] = 13420000
Life14$Population[42] = 11310000
```


```{r linear}
library(car)
#removing the country name
Life14 <- Life14[,-c(1,2)]
full.model <- lm(Life.expectancy~.,data=Life14)
vif(full.model)

#Observe residuals and outliers
par(mfrow=c(2,2))
plot(full.model, which =1) 

#Which obs are outliers according to Cooks D
cooksd <- cooks.distance(full.model)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")
plot(rstudent(full.model))
# Influence Plot
influencePlot(full.model,  main="Influence Plot")
#influencePlot(full.model, id.method="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )

#Review the observations that are pulled out as outliers


#Does a log of Life look better?
full.modellog <- lm(log(Life.expectancy)~.,data=Life14)
par(mfrow=c(2,2))
plot(full.modellog)

```

#Based off of reviewing the VIFs, there may be some multicollinearity for the following fields:
infant.deaths = 404
percentage.expenditure = 11.8
under.five.deaths = 325
GDP: 12
thinness 1-19: 12.8
thinness 5-9: 13
income.composition.of.resources = 11.4

#I don't see any indication by logging life that the residuals look more scattered in a random cloud. We will keep life expectancy as is for ease of interpretation. 

```{r variable selection}
library(leaps)
#Forward variable selection
reg.fwd = regsubsets(Life.expectancy~.,data=Life14,method="forward",nvmax=20)

par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:19,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:19,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:19,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

#Backward variable selection
reg.bwd = regsubsets(Life.expectancy~.,data=Life14,method="backward",nvmax=20)

par(mfrow=c(1,3))
bics<-summary(reg.bwd)$bic
plot(1:19,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.bwd)$adjr2
plot(1:19,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.bwd)$rss
plot(1:19,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

#Stepwise variable selection
reg.ex = regsubsets(Life.expectancy~.,data=Life14,method="exhaustive",nvmax=20)

par(mfrow=c(1,3))
bics<-summary(reg.ex)$bic
plot(1:19,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.ex)$adjr2
plot(1:19,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.ex)$rss
plot(1:19,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

#Looking at raw BIC
summary(reg.fwd)$bic
summary(reg.bwd)$bic
summary(reg.ex)$bic

```

Forward Selection seems to have the least parameter selection before accuracy decreases. 

```{r test train}
set.seed(123)
Set = sample(seq(1,dim(Life14)[1]),round(.50*dim(Life14)[1]),replace = FALSE)
#Training Set
Train_Life = Life14[Set,]
#Train_Life <- Train_Life[-1,]
#Test Set
Test_Life = Life14[-Set,]

#Check to see if it's pulled 50/50% 
#dim(Train_Life) #91   20
#dim(Test_Life) #91   20

reg.fwd=regsubsets(Life.expectancy~.,data=Train_Life,method="forward",nvmax=6)

#Really handy predict function
predict.regsubsets =function (object , newdata ,id ,...){
  form=as.formula (object$call [[2]])
  mat=model.matrix(form ,newdata )
  coefi=coef(object ,id=id)
  xvars=names(coefi)
  mat[,xvars]%*%coefi
}
testASE<-c()


#I cant figure out why this is throwing an error
##throwing an error from the NAs
for (i in 1:6){
  predictions<-predict.regsubsets(object=reg.fwd,newdata=Test_Life,id=i) 
  testASE[i]<-mean((Test_Life$Life.expectancy-predictions)^2)
}
par(mfrow=c(1,1))
plot(1:91,testASE,type="l",xlab="# of predictors",ylab="test vs train ASE",ylim=c(0.3,0.8))
index<-which(testASE==min(testASE))
points(index,testASE[index],col="red",pch=10)
rss<-summary(reg.fwd)$rss
lines(1:91,rss/100,lty=3,col="blue")  
```
